---
title: "LGCP Demo"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(spatstat)
set.seed(04222021)
```

#### Simulate and visualize LGCP

```{r}
m <- as.im(function(x, y){2 + 3*x + 3*y - 3* x^2}, W=owin())
X <- rLGCP("exp", m, var=0.1, scale =0.25)
plot(attr(X, "Lambda"), main = '')
points(X, col = rgb(0,0,0,.4), pch =16)
```

#### Fit LGCP

Note the plotting function separates the mean intensity and the error structure from the GP

```{r}
fitX <- kppm(X ~ x, "LGCP")
fitX
plot(fitX, what = 'intensity')

fitXY <- kppm(X ~ x + y, "LGCP")
fitXY
plot(fitXY, what = 'intensity')

fitX2Y <- kppm(X ~ polynom(x,2) + y , "LGCP")
fitX2Y
plot(fitX2Y, what = 'intensity')

```

#### BEI Dataset

Here is a KDE estimate of the surface intensity
```{r}
plot(density(bei))
points(bei, col = rgb(0,0,0,.1), pch =16)
```

### interaction with elevation and gradient
```{r}
fitBei <- ppm(bei ~ elev * grad , data=bei.extra)
fitBei
plot(fitBei, what = 'intensity')
```

### LGCP fit (kind of...)
```{r}
fitBei2 <- kppm(bei ~ elev * grad , "LGCP", data=bei.extra)
fitBei2
plot(fitBei2, what = 'intensity')
```

More comprehensive model fitting, may require approximate methods, and in particular Inverse Nested Laplace Approximations (INLA). See the following R packages for more details:

- `INLA`
- `inlabru`
- `lgcp`

