---
title: "Modeling Point Processes with Spatstat"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.align= 'center')
knitr::opts_chunk$set(fig.height= 4)
knitr::opts_chunk$set(fig.width = 6)
library(tidyverse)
library(gridExtra)
library(ggplot2)
library(smacpod)
library(spatstat)
set.seed(04082019)
```


### Model Fitting

The `ppm` function can be used for model fitting with a point process.

\vfill

```{r}
plot(bei)
```

\vfill

The `bei` dataset contains locations of trees in a tropical rain forest. The point pattern is clearly non-homogenous

\vfill
```{r}
plot(density.ppp(bei))
```
\vfill

\newpage

The pattern in the intensity of the trees may be related to elevation and the elevation gradient.

\vfill
```{r}
elev <- bei.extra$elev
grad <- bei.extra$grad
plot(elev)
plot(grad)
```

\newpage

The `ppm` function allows model fitting
\vfill
```{r}
tree.model <- ppm(bei ~ elev + grad);
tree.model
plot(tree.model)
```
\vfill

For more complicated models, `kppm` can be used for clustering behavior.

\vfill

### Marked Point Patterns
\vfill
Marked point process data contains meta data for each point. Rather than just $\boldsymbol{s}$, we have $(\boldsymbol{s},m)$.
\vfill

The marked information can either be categorical (multi-type) or continuous.

\vfill

The `lansing` data set contains locations of six types of trees.
```{r, fig.width=7, fig.height = 5}
plot(lansing, cols = 1:6)
```

```{r, fig.width=7, fig.height = 5}
plot(split(lansing))
```

\newpage

To analyze this data, consider the following model.
\vfill
```{r}
lansing.model <- ppm(lansing ~  marks - 1)
lansing.model
```

\vfill

In contrast with this model, we can also include 

```{r}
lansing.model2 <- ppm(lansing ~  marks * polynom(x,y,3))
#lansing.model2
plot(lansing.model2)
```
\vfill

Similarly continuous marked data can be included as a predictor i the `ppm` framework, potentially with interactions with spatially referenced data.

\vfill

Marked point process data can also be used for spatial-temporal point patterns, where the year corresponds to the mark.
\vfill


\newpage

### More advanced point pattern models

##### Cluster processes
Clustering is not well defined. In general the idea is that the point distances are shorter than expected. However, there "is a fundamental ambiguity between heterogeneity and clustering" (Diggle 2007).
\vfill

__Neyman-Scott Process__: This is a two stage process. 

\vfill
1. Generate parents 
\vfill
2. For each parent, generate a set of offspring
\vfill

__The shot noise processes__ are variations on the Neyman-Scott process, also with a two stage process.

\vfill

__Strauss Process__: contains a term that allows repulsion by adjusting the intensity in a vicinity of an existing point. The "hardcore" process will make the intensity 0 for any pair of points less than a specified distance $d_0$.
\vfill

